---
title: "RAProject"
author: "Mingzhi Ye"
date: "2022/10/10"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r,include = FALSE}
library(VDJdive)
library(tidyverse)
require(SingleCellExperiment)
library(r2r)



# metadata(sce)$clonoStats
# clonoAbundance(sce)
# clonoFrequency(sce)
# EMstats <- clonoStats(sce, method = "EM")
# 
# EMstats2=metadata(EMstats)$clonoStats
# barVDJ(EMstats2, title = "contigs", legend = TRUE)
# abundanceVDJ(EMstats2)
```

```{r}
sce0 <- readRDS('toyTCRdata.rds')
sce <- clonoStats(sce0, method = 'EM')

#A<-sce0$contigs[sce0$contigs[, 'chain'] %in% c('TRA','TRB')]
for (i in 1:length(sce0$contigs)){
  #sce0$contigs[[i]]  "DFrame"  "S4Vectors"
  #sce0$contigs[[i]]<-sce0$contigs[[i]] %>% filter(chain %in% c("TRA","TRB"))
  #sce0$contigs[[i]]<-sce0$contigs[[i]][filter(sce0$contigs[[i]]$chain %in% c("TRA","TRB")),]
  sce0$contigs[[i]]<-sce0$contigs[[i]][order(sce0$contigs[[i]]$chain),]
  
}
```

# Generate sample pool
```{r}
OriginalSamplePool=DataFrameList()
OriginalPoolSize=1
for (i in 1:length(sce0$contigs)){
  if(length(sce0$contigs[[i]]$chain)==2 & sce0$contigs[[i]]$chain[1]=='TRA' & sce0$contigs[[i]]$chain[2]=='TRB'){
    
    OriginalSamplePool[[OriginalPoolSize]]=sce0$contigs[[i]]
    OriginalPoolSize=OriginalPoolSize+1
  }
  
}
OriginalPoolSize=OriginalPoolSize-1


# Assignment different possiblity for each clonotype

assigned_num=rpois(OriginalPoolSize, 1)
assigned_num=assigned_num+1
assigned_index=c()
cursor=1
for (i in 1:length(assigned_num)){
  for(j in 1:assigned_num[i]){
    assigned_index[cursor]=i
    cursor=cursor+1
  }
}
samplePool=DataFrameList()
poolSize=1
for (i in 1:length(assigned_index)){
  
    
    samplePool[[poolSize]]=OriginalSamplePool[[assigned_index[i]]]
    poolSize=poolSize+1
  
  
}
poolSize=poolSize-1

```

# Functions for generate a sample
```{r}
getClonotype<-function(clonotype_index, numberA, numberB, errorProb = .01){
    result=DataFrame()
    indexInPool=sample.int(poolSize, 3, replace=FALSE)

    if (numberA>=1){
        if(runif(1) < errorProb){
          result=rbind(result,samplePool[[indexInPool[2]]][1,])
        }else{
          result=rbind(result,samplePool[[indexInPool[1]]][1,])
        }
        numberA=numberA-1
    }
    if (numberB>=1){
        if(runif(1) < errorProb){
          result=rbind(result,samplePool[[indexInPool[3]]][2,])
        }else{
          result=rbind(result,samplePool[[indexInPool[1]]][2,])
        }

        numberB=numberB-1
    }
    while(numberA>0 | numberB>0){
        indexInPool=sample.int(poolSize, 1, replace=TRUE)
        if (numberA>=1){
            result=rbind(result,samplePool[[indexInPool]][1,])
            numberA=numberA-1
        }
        else{
            result=rbind(result,samplePool[[indexInPool]][2,])
            numberB=numberB-1
        }
    }
    return(result)
    
}
```

# Use the functions to generate a sample with size 50

However, function ‘clonoStats’ is not compatible with SimpleDFrameList, how to transform SimpleDFrameList to CompressedSplitDFrameList
```{r}
getSample <- function(samplesize){
    sample=DataFrame()
    DistributeTRA=sum(sce0$contigs[,'chain']=='TRA')
    DistributeTRB=sum(sce0$contigs[,'chain']=='TRB')
    # table(DistributeTRA,DistributeTRB)
    IndexnAnBMap=cbind(DistributeTRA,DistributeTRB)
    RandomIntegers <- sample(1:poolSize, samplesize, replace=T)

    barcode <- paste('cell', 1:samplesize)
    
    
    sample <- lapply(1:samplesize, function(clonotype_index){
        getClonotype(clonotype_index, IndexnAnBMap[RandomIntegers[clonotype_index],1], IndexnAnBMap[RandomIntegers[clonotype_index],2])
    })
    sample <- SplitDataFrameList(sample)
    sample[,'barcode'] <- barcode
    sample[,'sample'] <- 'sim'
    names(sample) <- barcode
 
    # sample1<-as(sample,'SplitDFrameList')
    return(sample)
    
}
# samplelist=getSample(SampleSize)
# EMpredicted <- clonoStats(samplelist, method = 'EM')
# UNpredicted <- clonoStats(samplelist, method = 'unique')



```


# Generate hashmap. Key is clonotype in Truth(sample pool), Value is percentage of the clonotype in Truth

```{r}

TruthPercentageMap=hashmap()

for(i in 1:poolSize){
  
  key=paste(samplePool[[i]]$cdr3[1],samplePool[[i]]$cdr3[2],sep=' ')
  
  if(is.null(TruthPercentageMap[[key]])){
    TruthPercentageMap[[key]]<-1/poolSize
  }
  else{
    TruthPercentageMap[[key]]<-TruthPercentageMap[[key]]+1/poolSize
  }
  
}



```

# Get Variation Distance
```{r}
getDistance<-function(predicted){

  clonotypes=clonoNames(predicted)
  abundance=clonoAbundance(predicted)
  SumAbundance=sum(abundance)
  distance=0
  seen <- hashmap()
  for(i in 1:length(abundance)){
    if(is.null(TruthPercentageMap[[clonotypes[i]]])){
      distance=distance+abundance[i]/SumAbundance
    }
    else{
      seen[[clonotypes[i]]]<-TruthPercentageMap[[clonotypes[i]]]
      distance=distance+abs(abundance[i]/SumAbundance-TruthPercentageMap[[clonotypes[i]]])
    }
  }
  
  distance=distance+Reduce("+", values(TruthPercentageMap)) -Reduce("+", values(seen)) 
  return(distance)
}
```

# Calculate distances for samples

```{r,message = FALSE,warning = FALSE}
DistanceListFromEM=c()
DistanceListFromUN=c()
for(i in 1:10){
  samplelist=getSample(1000)
  EMpredicted <- clonoStats(samplelist, method = 'EM')
  distance1=getDistance(EMpredicted)
  DistanceListFromEM[i]=distance1
  UNpredicted <- clonoStats(samplelist, method = 'unique')
  distance2=getDistance(UNpredicted)
  DistanceListFromUN[i]=distance2
  print(i)
}

```


# Visualize
```{r}
boxplot(DistanceListFromEM)
hist(DistanceListFromEM)
boxplot(DistanceListFromUN)
hist(DistanceListFromUN)
print(median(DistanceListFromEM))
print(median(DistanceListFromUN))
print(mean(DistanceListFromEM))
print(mean(DistanceListFromUN))
```