---
title: "RAProject"
author: "Mingzhi Ye"
date: "2022/10/10"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r,include = FALSE}
library(VDJdive)
library(tidyverse)
require(SingleCellExperiment)
library(r2r)



# metadata(sce)$clonoStats
# clonoAbundance(sce)
# clonoFrequency(sce)
# EMstats <- clonoStats(sce, method = "EM")
# 
# EMstats2=metadata(EMstats)$clonoStats
# barVDJ(EMstats2, title = "contigs", legend = TRUE)
# abundanceVDJ(EMstats2)
```

```{r}
sce0 <- readRDS('toyTCRdata.rds')
sce <- clonoStats(sce0, method = 'EM')

#A<-sce0$contigs[sce0$contigs[, 'chain'] %in% c('TRA','TRB')]
for (i in 1:length(sce0$contigs)){
  #sce0$contigs[[i]]  "DFrame"  "S4Vectors"
  #sce0$contigs[[i]]<-sce0$contigs[[i]] %>% filter(chain %in% c("TRA","TRB"))
  #sce0$contigs[[i]]<-sce0$contigs[[i]][filter(sce0$contigs[[i]]$chain %in% c("TRA","TRB")),]
  sce0$contigs[[i]]<-sce0$contigs[[i]][order(sce0$contigs[[i]]$chain),]
  
}
```

# Generate sample pool
```{r}
samplePool=DataFrameList()
poolSize=1
for (i in 1:length(sce0$contigs)){
  if(length(sce0$contigs[[i]]$chain)==2 & sce0$contigs[[i]]$chain[1]=='TRA' & sce0$contigs[[i]]$chain[2]=='TRB'){
    
    samplePool[[poolSize]]=sce0$contigs[[i]]
    poolSize=poolSize+1
  }
  
}
poolSize=poolSize-1
```

# Functions for generate a sample
```{r}
getPoisson<-function(){
  a=-1
  while(a<0 | a>4){
    a=rpois(1, 1)
  }
  return(a)
}
getClonotype<-function(clonotype_index){
  numberA=getPoisson()
  numberB=getPoisson()
  result=DataFrame()
  indexInPool=sample.int(poolSize, 1, replace=TRUE)
  if (numberA>=1){
    result=rbind(result,samplePool[[indexInPool]][1,])
    numberA=numberA-1
  }
  if (numberB>=1){
    result=rbind(result,samplePool[[indexInPool]][2,])
    numberB=numberB-1
  }
  while(numberA>0 | numberB>0){
    indexInPool=sample.int(poolSize, 1, replace=TRUE)
    if (numberA>=1){
      result=rbind(result,samplePool[[indexInPool]][1,])
      numberA=numberA-1
    }
    else{
      result=rbind(result,samplePool[[indexInPool]][2,])
      numberB=numberB-1
    }
  }
  if(length(result$barcode)){
    new_barcode=paste(result$barcode[1],clonotype_index,sep=' ')
    for(i in 1:length(result$barcode)){
      #result$barcode[i]=result$barcode[1]
      result$barcode[i]=new_barcode
      result$sample[i]='sim'
    }
  }
  return(result)

}
```

# Use the functions to generate a sample with size 50

However, function ‘clonoStats’ is not compatible with SimpleDFrameList, how to transform SimpleDFrameList to CompressedSplitDFrameList
```{r}
getSample<-function(samplesize){
  sample=DataFrame()
  for(clonotype_index in 1:samplesize){
    sample=rbind(sample,getClonotype(clonotype_index))
  }
  samplelist<-split(sample,sample$barcode)
  # sample1<-as(sample,'SplitDFrameList')
  return(samplelist)
  
}
# samplelist=getSample(SampleSize)
# EMpredicted <- clonoStats(samplelist, method = 'EM')
# UNpredicted <- clonoStats(samplelist, method = 'unique')



```


# Generate hashmap. Key is clonotype in Truth(sample pool), Value is percentage of the clonotype in Truth

```{r}

TruthPercentageMap=hashmap()

for(i in 1:poolSize){
  
  key=paste(samplePool[[i]]$cdr3[1],samplePool[[i]]$cdr3[2],sep=' ')
  
  if(is.null(TruthPercentageMap[[key]])){
    TruthPercentageMap[[key]]<-1/poolSize
  }
  else{
    TruthPercentageMap[[key]]<-TruthPercentageMap[[key]]+1/poolSize
  }
  
}



```

# Get Variation Distance
```{r}
getDistance<-function(predicted){

  clonotypes=clonoNames(predicted)
  abundance=clonoAbundance(predicted)
  SumAbundance=sum(abundance)
  distance=0
  seen <- hashmap()
  for(i in 1:length(abundance)){
    if(is.null(TruthPercentageMap[[clonotypes[i]]])){
      distance=distance+abundance[i]/SumAbundance
    }
    else{
      seen[[clonotypes[i]]]<-TruthPercentageMap[[clonotypes[i]]]
      distance=distance+abs(abundance[i]/SumAbundance-TruthPercentageMap[[clonotypes[i]]])
    }
  }
  
  distance=distance+Reduce("+", values(TruthPercentageMap)) -Reduce("+", values(seen)) 
  return(distance)
}
```

# Calculate distances for samples

```{r,message = FALSE,warning = FALSE}
DistanceListFromEM=c()
DistanceListFromUN=c()
for(i in 1:50){
  samplelist=getSample(2000)
  EMpredicted <- clonoStats(samplelist, method = 'EM')
  distance1=getDistance(EMpredicted)
  DistanceListFromEM[i]=distance1
  UNpredicted <- clonoStats(samplelist, method = 'unique')
  distance2=getDistance(UNpredicted)
  DistanceListFromUN[i]=distance2
  print(i)
}

```


# Visualize
```{r}
boxplot(DistanceListFromEM)
hist(DistanceListFromEM)
boxplot(DistanceListFromUN)
hist(DistanceListFromUN)
print(median(DistanceListFromEM))
print(median(DistanceListFromUN))
print(mean(DistanceListFromEM))
print(mean(DistanceListFromUN))
```